---
title: "Trustworthiness assessment — <FIRST AUTHOR> (<YEAR>)"
subtitle: "<full citation + DOI>"
author: "Assessment prepared by <you>"
date: today
format:
  html:
    code-fold: true
    code-tools: true
    toc: true
    toc-float: true
    theme: flatly
    highlight-style: haddock
    embed-resources: true
editor_options:
  chunk_output_type: console
---

<!--
============================================================================
HOW TO USE — this is a SCAFFOLD, not a finished analysis.
  1. Copy this file into the paper's repo as
     code/<author>_<year>_trustworthiness_assessment.qmd, and copy the skill's
     scripts/helpers.R alongside it (this template does source("helpers.R")).
  2. Read references/forensic-method.md FIRST — method + stance. Re-read the
     stance before judging anything (burden on authors; no intent; think
     stupider; NO FALSE POSITIVES).
  3. Hand-key every reported statistic into the data objects below. Mark each
     scale integer vs non-integer (GRIM only applies to integer totals).
  4. Work the steps IN ORDER; delete steps that don't apply but say so. The
     deliverable is the written verdict at the foot, not the tables.
  R API notes (scrutiny digits_x gotcha, recalc, escalc): references/r-cookbook.md
============================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(scipen = 999)
```

```{r libraries}
library(tidyverse)
library(metafor)     # escalc(): SMD / effect sizes you derive
library(scrutiny)    # GRIM / GRIMMER
library(statcheck)   # recompute p from reported t/F/r/chi2 text
library(recalc)      # baseline p recalculation (install from ~/git/recalc)
library(knitr)
library(kableExtra)

source("helpers.R")  # skill glue: grim_n_profile, implied_sum, recalc_baseline_t/_chisq,
                     # prepost_r_plausibility (copy alongside this file).
                     # Stats engines called directly: scrutiny::grim*, recalc::* .
```

# Step 0 — Load-bearing claim

<!-- ONE sentence: the result the contribution rests on. Is it the pre-registered
PRIMARY outcome, or a secondary/post-hoc result promoted to the headline? -->

> *Load-bearing claim goes here.*

# Step 0.5 — Public record

<!-- Retraction / expression of concern on THIS paper; the TEAM's track record;
TRIAL REGISTRATION compared field-by-field to the paper. Tag [GOVERNANCE]. -->

```{r public-record}
# registration_check <- tribble(
#   ~field, ~registry, ~paper,
#   "Allocation", "...", "...",
#   "Masking",    "...", "...",
#   "Enrollment", "...", "...",
# )
# registration_check |> kable() |> kable_classic(full_width = FALSE)
```

# Extract data

<!-- Hand-key reported stats. Keep group/time/n explicit. `integer` flags whether
the score is an integer total (GRIM-eligible). Zung SDS/SAS index = raw x 1.25
are NOT integer-grained.

PRECISION RULE (non-negotiable). Key EVERY estimate at its full reported
precision, INCLUDING TRAILING ZEROS, and record the decimal count for each in an
explicit `*_digits` column. R drops trailing zeros from numerics (you type 13.50,
R stores 13.5), so the digit count — not the keyed number — is what carries the
precision into GRIM and the recalc intervals. NEVER default or infer digits: a
mean printed "13.50" is 2 dp, "0.050" is 3 dp, and getting this wrong flips
GRIM/recalc verdicts. Capture mean_digits / sd_digits / p_digits as you extract,
straight from how the paper prints each value. -->

```{r extract-data}
# Record reported precision per value (incl. trailing zeros) in *_digits columns.
# dat <- tribble(
#   ~variable, ~integer, ~condition, ~time, ~n, ~mean, ~mean_digits, ~sd, ~sd_digits,
#   ...
# )
# anova_tab <- tribble(~variable, ~F, ~f_digits, ~eta_reported, ~eta_digits, ...)
```

# Step 1 — Numerical sanity check

## F ↔ partial η² (and the N it implies)

```{r eta-F}
# f_digits / eta_digits = the reported decimal places (incl. trailing zeros):
# recalc::recalc_partial_eta_from_f(f = <F>, df_effect = 1, df_error = <df>,
#   f_digits = <F dp>, eta = <reported eta^2>, eta_digits = <eta dp>)   # $consistent
# (df_error = N - #groups also cross-checks the N every test assumes.)
```

## Correlations r → t → p

```{r corr}
# tibble(r = c(...), n = <n>) |>
#   mutate(t = r*sqrt((n-2)/(1-r^2)), p = round(2*pt(-abs(t), n-2), 4)) |>
#   kable() |> kable_classic(full_width = FALSE)
```

## Participant-flow check

<!-- Do all the n's add up across CONSORT diagram, text, and tables? Does the
test df imply a single N? Hold this against Step 2 (GRIM alternative-n). -->

## Baseline p-value recalculation (`recalc`)

<!-- Each baseline row reports m/sd/n (or counts) AND a p → recompute each p.
A reported p OUTSIDE its recomputed hull = [IMPOSSIBLE]. If all reproduce, SAY SO
(clean numerical-sanity result). The distribution of trusted p's is a Carlisle
check in Step 3 — weak with few variables. -->

```{r recalc-baseline}
# Every row carries its reported decimal places (incl. trailing zeros): no defaults.
# baseline_cont <- tribble(
#   ~variable, ~m1, ~sd1, ~m2, ~sd2, ~p, ~m_digits, ~sd_digits, ~p_digits,
#   "HAMD", 24.71, 4.137, 24.81, 3.774, .921,  2, 3, 3,
#   ...
# )
# recalc_baseline_t(baseline_cont, n1 = <n1>, n2 = <n2>) |>
#   kable(caption = "Reported vs recomputed baseline p (multiverse hull).") |>
#   kable_classic(full_width = FALSE)
# # categorical row (state p_digits, incl. trailing zeros):
# recalc_baseline_chisq(matrix(c(a,b,c,d), nrow = 2, byrow = TRUE), p = .082, p_digits = 3)
```

# Step 2 — Possibility: GRIM / GRIMMER

<!-- KEY PATTERN: run GRIM SEPARATELY by group x timepoint. Asymmetry (one block
consistent, a sibling block not, at the same reported n) is the signal. Then
search alternative n on the failing block. Tag failures [IMPOSSIBLE] with the
arithmetic; distinguish rounding artifacts from true contradictions. Also: bounds
(Popoviciu SD<=range/2, Bhatia-Davis, |r|<=1, estimate inside its CI, PSD
correlation matrix) and coherence (part-whole; bootstrap-CI-vs-p). -->

```{r grim-block}
# GRIM is scrutiny's; pass the per-value reported precision (incl. trailing zeros)
# from the mean_digits column — scrutiny::grim's digits_x is vectorised.
# grim_res <- dat |> filter(integer) |>
#   mutate(GRIM = scrutiny::grim(mean, n, digits_x = mean_digits)) |>
# grim_res |> group_by(condition, time) |>
#   summarise(pass = sum(GRIM), total = n(), .groups = "drop") |>
#   kable() |> kable_classic(full_width = FALSE)
```

```{r grim-altn}
# failing <- dat |> filter(integer, condition == "<arm>", time == "<time>")
# prof <- grim_n_profile(failing$mean, n_range = 15:40, digits = failing$mean_digits)
# ggplot(prof, aes(n, consistent)) + geom_col() + theme_minimal()
# map_dfr(failing$mean, implied_sum, n = <peak_n>)   # residual ~0 confirms
```

```{r grimmer}
# digits_x / digits_sd come from the reported precision (incl. trailing zeros):
# scrutiny::grimmer_map(dat |> filter(integer) |> transmute(x = mean, sd, n),
#                       digits_x = <mean dp>, digits_sd = <sd dp>) |>
#   select(x, sd, n, consistency)   # more assumption-laden; report cautiously
```

# Step 3 — Plausibility

<!-- Derive omitted effect sizes; test against the design ceiling and the field.
Watch the SE-vs-SD swap. For pre/post data, check change-score coherence (implied
pre-post r) below. Carlisle-style baseline distribution (weak with few vars).
Tag [IMPLAUSIBLE] / [IMPOSSIBLE]. -->

```{r effect-sizes}
# escalc(measure = "SMD", m1i=, sd1i=, n1i=, m2i=, sd2i=, n2i=) |> summary()
```

```{r implied-r}
# Change-score coherence -> implied within-person pre-post r (recalc >= 0.6).
# recalc returns the recalculated [lower, upper] interval (rounding-propagated).
# All *_digits = the reported decimal places (incl. trailing zeros); never default.
# (A) when a change SD is reported (or an independent t was run on change scores):
# rc <- recalc::recalc_prepost_r(sd_pre = , sd_post = , sd_change = ,
#         sd_pre_digits = <dp>, sd_post_digits = <dp>, sd_change_digits = <dp>)
# (B) when only a 2x2 (group x time) RM-ANOVA interaction F is given:
# rc <- recalc::recalc_prepost_r_from_f(f = <F>,
#         m1b, sd1b, m1p, sd1p, n1, m2b, sd2b, m2p, sd2p, n2,
#         f_digits = <dp>, m_digits = <dp>, sd_digits = <dp>)
#
# POSSIBILITY (recalc): SDs can't coexist when the interval is wholly outside [-1,1]:
#   impossible <- rc$recalculated_lower > 1 || rc$recalculated_upper < -1   # [IMPOSSIBLE]
# PLAUSIBILITY (skill): for a possible r, label the point estimate:
#   prepost_r_plausibility((rc$recalculated_lower + rc$recalculated_upper)/2)
#   # implausible_negative / implausibly_high -> [IMPLAUSIBLE]; bands follow `within`.
# rc |> kable(caption = "Implied pre-post r (recalculated interval).") |> kable_classic(full_width = FALSE)
```

# Step 4 — Misalignment

<!-- measure<->construct; test<->inference (causal language from observational
data; efficacy without a controlled post comparison; change-change correlation as
mechanism; ANCOVA on non-equivalent groups; cluster-then-test; Table-2 fallacy);
claim<->evidence (abstract vs tables; general claim on a narrow effect; estimand
never defined). Tag [MISALIGNMENT]. -->

> *Misalignment findings (prose), each tagged and argued.*

# Step 5 — Conduct, governance, transparency

<!-- ethics/registration timing & consistency (Step 0.5); feasibility; data
availability; duplicate text/figures; consistency across the team's companion
papers. Tag [GOVERNANCE]. -->

> *Governance findings (prose), each tagged.*

# Output — verdict

<!--
1. CLASSIFICATION (one): trustworthy as reported / concerns pending clarification
   (specify what resolves them) / not trustworthy as reported.
2. BOTTOM LINE (1-2 sentences).
3. NUMERICAL SANITY CHECK (what held).
4. FINDINGS ORDERED BY SEVERITY, each tagged, arithmetic shown.
5. WHAT I COULD NOT CHECK.
Close with the reader caution: a starting point, not a verdict to act on
unchecked; re-verify arithmetic; describe issues, do not accuse.
-->
