---
title: "Trustworthiness assessment — <FIRST AUTHOR> (<YEAR>)"
subtitle: "<full citation + DOI>"
author: "Assessment prepared by <you>"
date: today
format:
  html:
    code-fold: true
    code-tools: true
    toc: true
    toc-float: true
    theme: flatly
    highlight-style: haddock
    embed-resources: true
editor_options:
  chunk_output_type: console
---

<!--
============================================================================
HOW TO USE — this is a SCAFFOLD, not a finished analysis.
  1. Copy this file into the paper's repo as
     code/<author>_<year>_trustworthiness_assessment.qmd, and copy the skill's
     scripts/helpers.R alongside it (this template does source("helpers.R")).
  2. Read references/forensic-method.md FIRST — method + stance. Re-read the
     stance before judging anything (burden on authors; no intent; think
     stupider; NO FALSE POSITIVES).
  3. Hand-key every reported statistic into the data objects below. Mark each
     scale integer vs non-integer (GRIM only applies to integer totals).
  4. Work the steps IN ORDER; delete steps that don't apply but say so. The
     deliverable is the written verdict at the foot, not the tables.
  R API notes (scrutiny digits_x gotcha, recalc, escalc): references/r-cookbook.md
============================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(scipen = 999)
```

```{r libraries}
library(tidyverse)
library(metafor)     # escalc(): SMD / effect sizes you derive
library(scrutiny)    # GRIM / GRIMMER
library(statcheck)   # recompute p from reported t/F/r/chi2 text
library(recalc)      # baseline p recalculation (install from ~/git/recalc)
library(knitr)
library(kableExtra)

source("helpers.R")  # grim_consistent, grim_n_profile, recalc_baseline_t,
                     # implied_prepost_r, eta_from_F, ...  (copy alongside this file)
```

# Step 0 — Load-bearing claim

<!-- ONE sentence: the result the contribution rests on. Is it the pre-registered
PRIMARY outcome, or a secondary/post-hoc result promoted to the headline? -->

> *Load-bearing claim goes here.*

# Step 0.5 — Public record

<!-- Retraction / expression of concern on THIS paper; the TEAM's track record;
TRIAL REGISTRATION compared field-by-field to the paper. Tag [GOVERNANCE]. -->

```{r public-record}
# registration_check <- tribble(
#   ~field, ~registry, ~paper,
#   "Allocation", "...", "...",
#   "Masking",    "...", "...",
#   "Enrollment", "...", "...",
# )
# registration_check |> kable() |> kable_classic(full_width = FALSE)
```

# Extract data

<!-- Hand-key reported stats. Keep group/time/n explicit. `integer` flags whether
the score is an integer total (GRIM-eligible). Zung SDS/SAS index = raw x 1.25
are NOT integer-grained. -->

```{r extract-data}
# dat <- tribble(
#   ~variable, ~integer, ~condition, ~time, ~n, ~mean, ~sd,
#   ...
# )
# anova_tab <- tribble(~variable, ~F, ~eta_reported, ...)   # if RM-ANOVA reported
```

# Step 1 — Numerical sanity check

## F ↔ partial η² (and the N it implies)

```{r eta-F}
# anova_tab |>
#   mutate(eta_calc = round(eta_from_F(F, df_error = <df>), 3),
#          diff = eta_calc - eta_reported) |>
#   kable() |> kable_classic(full_width = FALSE)
```

## Correlations r → t → p

```{r corr}
# tibble(r = c(...), n = <n>) |>
#   mutate(t = r*sqrt((n-2)/(1-r^2)), p = round(2*pt(-abs(t), n-2), 4)) |>
#   kable() |> kable_classic(full_width = FALSE)
```

## Participant-flow check

<!-- Do all the n's add up across CONSORT diagram, text, and tables? Does the
test df imply a single N? Hold this against Step 2 (GRIM alternative-n). -->

## Baseline p-value recalculation (`recalc`)

<!-- Each baseline row reports m/sd/n (or counts) AND a p → recompute each p.
A reported p OUTSIDE its recomputed hull = [IMPOSSIBLE]. If all reproduce, SAY SO
(clean numerical-sanity result). The distribution of trusted p's is a Carlisle
check in Step 3 — weak with few variables. -->

```{r recalc-baseline}
# baseline_cont <- tribble(
#   ~variable, ~m1, ~sd1, ~m2, ~sd2, ~p,
#   "HAMD", 24.71, 4.137, 24.81, 3.774, .921,
#   ...
# )
# recalc_baseline_t(baseline_cont, n1 = <n1>, n2 = <n2>) |>
#   kable(caption = "Reported vs recomputed baseline p (multiverse hull).") |>
#   kable_classic(full_width = FALSE)
# # categorical row:
# recalc_baseline_chisq(matrix(c(a,b,c,d), nrow = 2, byrow = TRUE), p = .082)
```

# Step 2 — Possibility: GRIM / GRIMMER

<!-- KEY PATTERN: run GRIM SEPARATELY by group x timepoint. Asymmetry (one block
consistent, a sibling block not, at the same reported n) is the signal. Then
search alternative n on the failing block. Tag failures [IMPOSSIBLE] with the
arithmetic; distinguish rounding artifacts from true contradictions. Also: bounds
(Popoviciu SD<=range/2, Bhatia-Davis, |r|<=1, estimate inside its CI, PSD
correlation matrix) and coherence (part-whole; bootstrap-CI-vs-p). -->

```{r grim-block}
# grim_res <- dat |> filter(integer) |>
#   rowwise() |> mutate(GRIM = grim_consistent(mean, n)) |> ungroup()
# grim_res |> group_by(condition, time) |>
#   summarise(pass = sum(GRIM), total = n(), .groups = "drop") |>
#   kable() |> kable_classic(full_width = FALSE)
```

```{r grim-altn}
# failing <- dat |> filter(integer, condition == "<arm>", time == "<time>")
# prof <- grim_n_profile(failing$mean, n_range = 15:40)
# ggplot(prof, aes(n, consistent)) + geom_col() + theme_minimal()
# map_dfr(failing$mean, implied_sum, n = <peak_n>)   # residual ~0 confirms
```

```{r grimmer}
# scrutiny::grimmer_map(dat |> filter(integer) |> transmute(x = mean, sd, n),
#                       digits_x = 2, digits_sd = 3) |>
#   select(x, sd, n, consistency)   # more assumption-laden; report cautiously
```

# Step 3 — Plausibility

<!-- Derive omitted effect sizes; test against the design ceiling and the field.
Watch the SE-vs-SD swap. For 2x2 RM-ANOVA also interrogate the interaction F via
the implied change-SD / pre-post r. Carlisle-style baseline distribution (weak
with few vars). Tag [IMPLAUSIBLE]. -->

```{r effect-sizes}
# escalc(measure = "SMD", m1i=, sd1i=, n1i=, m2i=, sd2i=, n2i=) |> summary()
```

```{r implied-r}
# implied_prepost_r(F_int = <interaction F>,
#   m1b, s1b, m1p, s1p, n1, m2b, s2b, m2p, s2p, n2) |>
#   kable(caption = "Implied change-score SD & pre-post r from the interaction F.") |>
#   kable_classic(full_width = FALSE)
```

# Step 4 — Misalignment

<!-- measure<->construct; test<->inference (causal language from observational
data; efficacy without a controlled post comparison; change-change correlation as
mechanism; ANCOVA on non-equivalent groups; cluster-then-test; Table-2 fallacy);
claim<->evidence (abstract vs tables; general claim on a narrow effect; estimand
never defined). Tag [MISALIGNMENT]. -->

> *Misalignment findings (prose), each tagged and argued.*

# Step 5 — Conduct, governance, transparency

<!-- ethics/registration timing & consistency (Step 0.5); feasibility; data
availability; duplicate text/figures; consistency across the team's companion
papers. Tag [GOVERNANCE]. -->

> *Governance findings (prose), each tagged.*

# Output — verdict

<!--
1. CLASSIFICATION (one): trustworthy as reported / concerns pending clarification
   (specify what resolves them) / not trustworthy as reported.
2. BOTTOM LINE (1-2 sentences).
3. NUMERICAL SANITY CHECK (what held).
4. FINDINGS ORDERED BY SEVERITY, each tagged, arithmetic shown.
5. WHAT I COULD NOT CHECK.
Close with the reader caution: a starting point, not a verdict to act on
unchecked; re-verify arithmetic; describe issues, do not accuse.
-->
